---
title: "Data from open sources"
author: '@revaluation'
date: "6/23/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)

```

# Global Power Plant Database  

**Source:**  <https://datasets.wri.org/dataset/globalpowerplantdatabase>  
Global Energy Observatory, Google, KTH Royal Institute of Technology in Stockholm, Enipedia, World Resources Institute. 2018. Global Power Plant Database. Published on Resource Watch and Google Earth Engine; http://resourcewatch.org/ https://earthengine.google.com/  

**License:**  
Creative Commons Attribution 4.0 International License. Full license text available at (Creative Commons Attribution 4.0)[http://www.wri.org/publications/permissions-licensing]  

**Metadata**  
Project: Climate Data and Tools
Page Last Updated: December 2, 2019

```{r}
dir_wri <- "data/WRI"
if (!dir.exists(dir_wri)) dir.create(dir_wri, recursive = TRUE, showWarnings = FALSE)

if (!file.exists(file.path(dir_wri, "india_power_plants.RData"))) {
  www <- "https://wri-dataportal-prod.s3.amazonaws.com/c24/0ed/2e-1190-4d7e-b1da-c66b72e08858/globalpowerplantdatabasev120.zip"
  (fl <- file.path(dir_wri, basename(www)))
  
  dw <- download.file(www, destfile = fl)
  stopifnot(dw == 0) 
  unzip(zipfile = fl, exdir = dir_wri)
  unlink(fl)
  
  global_power_plant_database <- read.csv(file.path(dir_wri, "global_power_plant_database.csv")) %>%
    as_tibble()
  # save(global_power_plant_database, file = file.path(dir_wri, "global_power_plant_database.RData"))
  india_power_plants <- global_power_plant_database %>%
    filter(country == "IND")
  rm(global_power_plant_database)
  save(india_power_plants, file = file.path(dir_wri, "india_power_plants.RData"))
  
  pp <- select(india_power_plants, -country, - country_long)
  unique(pp$primary_fuel)
  
  (load("data/maps/india_adm1_light.RData"))
  plot(india_light_sp)
  points(pp$longitude, pp$latitude, pch = 16, col = "red")

  # Associating power plants with states
  # ?sp::over
  pp$region <- NA
  for (i in 1:dim(pp)[1]) {
    pp_sp <- SpatialPoints(cbind(pp$longitude[i], pp$latitude[i]), proj4string = india_light_sp@proj4string)
    pp_reg <- over(pp_sp, india_light_sp)$region
    if (length(pp_reg) != 1) warning("i = ", i, ", regions: ", pp_reg)
    if (is.na(pp_reg)) {
      pp_d <- rgeos::gDistance(pp_sp, india_light_sp, byid = T)
      pp_reg <- india_light_sp$region[which.min(pp_d)]
    }
    
    pp$region[i] <- pp_reg
    pp_reg <- NULL
  }
  ii <- is.na(pp$region); summary(ii)
  
  if (sum(ii) > 0) {
    pp[ii,]
    
    plot(india_light_sp, main = "Unidentified points")
    points(pp$longitude[ii], pp$latitude[ii], pch = 16, col = "red")
    
    # ?gBuffer # 
  }

  save(pp, file = file.path(dir_wri, "india_power_plants.RData"))

} else {
  load(file.path(dir_wri, "india_power_plants.RData"))
}

ggplot(data = india_light_df) + 
  geom_polygon(aes(x = long, y = lat, group = group), fill = "wheat", 
               colour = "white", alpha = 1, size = .5) + # aes fill = id, 
  coord_fixed(1.) +
  # guides(fill=FALSE) + # do this to leave off the color legend
  theme_void() + 
  labs(title = "India's power plants (Global Power Plant Database, WRI, 2019)")  +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  geom_point(data = pp, 
             aes(x = longitude, y = latitude, colour = primary_fuel, size = capacity_mw),
             alpha = .7) +
  guides(color = guide_legend(override.aes = list(size=5))) 
  # scale_color_brewer(palette = "Set2")

```

